---
title: "Deterministic Monte Carlo Mixing Model of Regional Groundwater Salinization"
author: "Rich Pauloo"
date: "7/4/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### All scripts, versions, data input files, and documentation for this model can be found in [this Github repository.](https://github.com/RichPau/Monte-Carlo-Mixing-Model)

## Set working directory and load data [^1]
##### C2VSim Groundwater, Land, and Root Zone Budgets are for the 40 year period from 10/31/1961-9/30/2001. [@Brush2013].
```{r set working directory and load data, warning = F, message = F}
# Groundwater budget (10/31/1961-9/30/2001)
GW = read.csv(file = "data/GW.csv", stringsAsFactors = FALSE, header = TRUE) 

# Land Budget budget (10/31/1961-9/30/2001)
LB = read.csv(file = "data/LB.csv", stringsAsFactors = FALSE, header = TRUE)

# Root Zone budget (10/31/1961-9/30/2001)
RZ = read.csv(file = "data/RZ.csv", stringsAsFactors = FALSE, header = TRUE) 

# Bring in boundary condition data from Tulare_well_stats.Rmd
library(dplyr)
boundary_dat <- readRDS("data/boundary_dat.rds")
bd <- boundary_dat %>% dplyr::select(x,y) %>% 
  mutate(x = abs(x) * 3.28084) %>% # convert m to ft for model
  arrange(x) %>% 
  as.matrix()
```
[^1]: Original C2VSim model files (R374) can be found at: <http://baydeltaoffice.water.ca.gov/modeling/hydrology/C2VSim/index_C2VSIM.cfm>

## Load Required Packages [^2]
```{r load packages, warning = F, message = F}
library(extrafont)
library(scales)
library(ggplot2)
library(grid)
library(gridExtra)
```
[^2]: if these packages are not already installed, use the function ```install.packages()``` to install the necessary packages

## Define Common Unit Conversion Functions
```{r Unit Conversion Functions, warning = F, message = F}
# converts acre-feet to liters
af_to_L = function(a){
  a * 43560 * 28.3168
}

# converts miligrams to million tons
mg_to_Mton = function(a){
  a / (1000*907185*1000000)
}
```

## Define Boundary Conditions 
##### Rock Water Interaction Coefficients dervied from groundwater salinity literature of the Tulare Basin [@Kang2016; @Williamson1989]. 
##### Groundwater velocity was computed with Z Budget and the C2VSim model, Version 3.02-CG (R374) [@Brush2013].

```{r boundary conditions, warning = F, message = F}
# Rock water interactions
Kang       = 3000/mean(c(1719,2205)) # (Kang, 2016) 3000 mg/L at 1719 ft. in Fresno co., 2205 ft. in Kern co.
Williamson = 2000/2500

# velocity field from C2VSim
dv12 = 0.76106 # Darcy velocity between centroids of layers 1-2 (ft/yr)
dv23 = -0.01772 # Darcy velocity between centroids of layers 2-3 (ft/yr)

# C2VSim Groundwater budget (units = AF)
DP     = sum(GW$DP) # deep percolation
Si     = sum(GW$Si) # beginning storage
Sf     = sum(GW$Sf) # ending storage
NDP    = sum(GW$NDP) # net deep percolation
Stream = sum(GW$Stream) # stream gain
R      = sum(GW$R) # recharge
Lake   = sum(GW$Lake) # lake gain
BI     = sum(GW$BI) # sum(GW$Si) boundary inflow
Sub    = sum(GW$Sub) # subsidence flow
P      = sum(GW$P) # GW pumping
SI     = sum(GW$Net.Subsurface.Inflow....) # subsurface inflow

# spatial/temporal dimensions
Dim = data.frame(subregion = seq(14,21,1), 
                 SA = c(670228.87, 904472.46, 302449.08, 372889.22, 
                        897090.81, 801419.62, 423713.29, 652847.11), # acre feet
                 thickness = c(1915.155, 1990.01, 1509.194, 1868.821, 
                               1904.244, 1715.117, 1969.524, 3244.262)) # feet: from C2VSim model

Tot_SA         = sum(Dim$SA) # total surface area (acres)
V              = Dim$SA * Dim$thickness # Volume (AF)
Tot_V          = af_to_L(sum(V)) # total volume (L)
SA_w_thickness = sum(Dim$SA/Tot_SA*Dim$thickness) # surface area weighted thickness

# time step length
t = 50 # years

# length of C2VSim simulation
l = 40 # years


####################################################################################
# salt concentration of natural SW and thus GWW from (Cismowski, 2006)
# within report search for: Annual Salt Load (thousand tons/year)
####################################################################################
sw <- data.frame(term = c("sac","yolo","sjr"),
                 ttons = c(1945, 405,922),
                 taf_yr = c(16953,2980,3082))

# concentration of delivered sw
sw$c <- sw$ttons / sw$taf_yr
sw$c <- sw$c * 9.072e+11 / 1.233e+9 # convert ktons to mg and thousand acre feet to L

# volume weighted concetration
sw <- sw %>% 
  mutate(pv  = taf_yr / (sum(sw$taf_yr)), # percent volume
         vwc = pv * c) # volume weighted concentration
# sum(sw$vwc) # final concentration of surface water input

# volume weighted concetration
sw <- sw %>% 
  mutate(pv  = taf_yr / (sum(sw$taf_yr)), # percent volume
         vwc = pv * c) # volume weighted concentration
gwc <- sum(sw$vwc) # final concentration of natural surface water and thus GW
####################################################################################

# C2VSim average annual fluxes from GW budget
annual_fluxes = data.frame(term = c("NDP", "Stream", "R", "Lake", "BI", "Sub", "P", "SI"),
                           "L/yr" = c(af_to_L(NDP/l), af_to_L(Stream/l), af_to_L(R/l), 
                                      af_to_L(Lake/l), af_to_L(BI/l), af_to_L(Sub/l), 
                                      af_to_L(P/l), af_to_L(SI/l)))
annual_fluxes$TDS = c(0,gwc,gwc,gwc,gwc,gwc,0,gwc) # TDS
annual_fluxes$mass_flux = annual_fluxes$L.yr * annual_fluxes$TDS # mass flux (mg/yr)

# Internal mass flux
internal_mf = (annual_fluxes[5,4] + annual_fluxes[6,4] + annual_fluxes[8,4])*t # mg/50 yr 

# Top down Mass Flux
top_mf = (annual_fluxes[2,4] + annual_fluxes[3,4] + annual_fluxes[4,4])*t # mg/50 yr 

# Root Zone Budget (includes Agriculture, Urban, and Native Vegetation)
I  = sum(RZ$AG_INF) + sum(RZ$URB_INF) + sum(RZ$NAT_INF) # Net infiltration
ET = sum(RZ$AG_ET) + sum(RZ$URB_ET) + sum(RZ$NAT_ET) # Net ET

# Percentages
pI   = NDP/I # percentage of infiltration that becomes NDP
pGWP = P/I # Percentage of groundwater pumping in Total Applied Water
pSW  = 1 - pGWP # percentage of surface water diversions and reuse in Total Applied Water

# Average Annual Surface Water TDS, volume, and mass flux
# TDS_SW = .38*((247.22540+254.26799)/2) + .62*(87.24187) # TDS of surface water

####################################################################################
# salt concentration of SW imports from (Cismowski, 2006)
# within report search for: Annual Salt Load (thousand tons/year)
####################################################################################

# period of record 1985 - 1994
sw <- data.frame(term = c("swp","cvp"),
                 ttons = c(1004, 900),
                 taf_yr = c(2471,2328))

# concentration of delivered sw
sw$c <- sw$ttons / sw$taf_yr
sw$c <- sw$c * 9.072e+11 / 1.233e+9 # convert ktons to mg and thousand acre feet to L

# volume weighted concetration
sw <- sw %>% 
  mutate(pv  = taf_yr / (sum(sw$taf_yr)), # percent volume
         vwc = pv * c) # volume weighted concentration
TDS_SW <- sum(sw$vwc) # final concentration of surface water input
####################################################################################

V_SW = (I - P) / l # Average Annual SW diversion volume (AF/yr)
SW_annual_mass = mg_to_Mton(af_to_L(V_SW) * TDS_SW) # Mton
```

## Number of Monte Carlo Samples
###### This number determines how many random samples are drawn from the input parameters modeled as distributions of a random variable. 1000 samples is standard.
```{r Number of Monte Carlo Samples, warning = F, message = F}
N = 1000
```

## Initialize 3D output arrays 
```{r Initialize 3D output arrays, warning = F, message = F}
TDS                  = array(0, dim=c(8,8,N)) # TDS array
layer_depths_array   = array(0, dim=c(8,1,N)) # layer depths array
layer_velocity_array = array(0, dim=c(8,1,N)) # layer velocity array
mass                 = array(0, dim=c(8,1,N)) # array of sum of masses per time step
TDS_GWP              = array(0, dim=c(8,1,N)) # array of TDS of pumped groundwater per time step
TDS_AW               = array(0, dim=c(8,1,N)) # array of TDS of applied water (SW+GW) per time step
EC                   = array(0, dim=c(8,1,N)) # array of evaporative concentration of NDP per time step (TDS)
NDP_mf               = array(0, dim=c(8,1,N)) # array of NDP mass flux per time step (mg/yr)
RWI_cont             = array(0, dim=c(8,1,N)) # array of Rock water interactions
lay_vol              = array(0, dim=c(8,1,N)) # array of layer volumes
```

## Compute all output  
```{r Compute all Output, warning = F, message = F}
set.seed(23545) # ensure replicability

for(k in 1:N){

  # Random Variables
  n    = runif(1, min = 0.25, max = 0.35) # porosity
  pa   = runif(1, min = 0.35, max = 0.50) # percent aquifer
  aVol = Tot_V * n * pa # saturated aquifer volume (L)
  RWI  = runif(1, min = Williamson, max = Kang)  # Rock water interactions (TDS/ft) 
  
  # velocity field from C2VSim
  v12    = 0.76106/n # linear pore velocity between centroids of layers 1-2 (ft/yr)
  v23    = -0.01772/n # linear pore velocity between centroids of layers 2-3 (ft/yr)
  vel    = c(v12,v23) # vector of velocities
  depths = c(-631.5,-2077) # C2VSim midpoints between layer centroids over which velcoity is calculated
  vel    = as.data.frame(cbind(vel,depths)) # bind vel and depth vectors into one data frame for analysis
  lm     = lm(vel$depths~vel$vel) # create a linear model between velocity and depth
  
  # vertical velocity profile
  v    = matrix(0,8,1) # initalize vector to hold computed velocities at various depths
  b    = matrix(0,8,1) # initalize vector to hold layer thicknesses, computed from groundwater velocity
  v[1] = (0 + -coef(lm)[[1]]) / coef(lm)[[2]] # compute first velocity
  b[1] = -(v[1] * t) # compute first layer thickness
  
  # compute all remaining layer velocities and thicknesses
  for(i in 1:7){
    v[i+1] = (sum(b) -coef(lm)[[1]]) / coef(lm)[[2]]
    b[i+1] = -(v[i+1] * t)
  }
  
  # Vertical Flux per layer (L/yr)
  Q = v * n * pa * Tot_SA * 43560 * 28.3168 # L/yr
  layer_fluxes = cbind(v, Q) # units of "ft/yr" and "L/yr" respectively
  
  # layer thickness (ft)
  b[8] = -(SA_w_thickness + sum(b[1:7]))
  pb   = abs(b)/SA_w_thickness # percentage of total thickness per layer
  
  # layer volumes (L)
  lay_vol[,1,k] = aVol * pb # saturated volume of all layers (L)
  

  
  # Calculate Pumping in each layer
  lay_pump = matrix(0,8,1)
  for(i in 1:7){
    lay_pump[i,1] = b[i]/sum(b) * annual_fluxes[7,2] # proportion pumping from each layer (L/yr)
  }
  

  
  # Calculate the layer depths
  layer_depths    = matrix(0,8)
  layer_depths[1] = abs(b[1])
  for(i in 2:8){
    layer_depths[i,1] = abs(layer_depths[i-1,1]) + abs(b[i]) # vector of layer depths (ft)
  }
  

  
  # Initalize TDS array with baseline TDS
  #t0 = layer_depths * RWI # inital TDS-depth profile from RWI
  
  bc_match <- matrix(0, nrow = 8, ncol = 1)
  for(i in 1:nrow(layer_depths)){
    temp  <- layer_depths[i,1] - bd[,1]
    index <- which.min(abs(temp))
    bc_match[i] = bd[index,2]
  }
  
  t0 = bc_match
  TDS[,1,k] = t0
  
  # RWI == FALSE 
  RWI_cont[,1,k] = 0
  
  # # RWI == TRUE
  # # Initalize TDS array with baseline TDS
  # t0 = layer_depths * RWI # inital TDS-depth profile from RWI
  # TDS[,1,k] = t0
  # # RWI
  # RWI_cont[,1,k] = t0

  
  # Record layer depth/velocity for all simulations
  layer_depths_array[,,k]   = layer_depths
  layer_velocity_array[,,k] = v
  

  
  # Compute TDS for all layers and times
  for(j in 1:7){
    
    # calculate mass flux of NDP
    mass[j,1,k]    = sum(TDS[1:7,j,k] * lay_vol[1:7,1,k]) # sum of mass in pumping layers 1-7 (mg)
    TDS_GWP[j,1,k] = mass[j,1,k]/sum(lay_vol[1:7,1,k]) # TDS of pumped GW
    TDS_AW[j,1,k]  = TDS_GWP[j,1,k] * pGWP + TDS_SW * pSW # TDS of applied water for irrigation
    EC[j,1,k]      = TDS_AW[j,1,k] * (1/pI) # TDS of NDP Evaporative concentration
    NDP_mf[j,1,k]  = EC[j,1,k] * annual_fluxes[1,2] # mass flux of NDP
    
    # TDS in layer 1
    TDS[1,j+1,k] = ((TDS[1,j,k]*lay_vol[1,1,k] + top_mf + NDP_mf[j,1,k]*t - TDS[1,j,k]*layer_fluxes[2,2]*t - 
                     TDS[1,j,k]*lay_pump[1,1]*t + internal_mf*pb[1]) / lay_vol[1,1,k]) + RWI_cont[1,1,k]
    
    # TDS in layers 2-7
    for(i in 2:7){
      TDS[i,j+1,k] = ((TDS[i,j,k]*lay_vol[i,1,k] + internal_mf*pb[i] + TDS[i-1,j,k]*layer_fluxes[i,2]*t -
                       TDS[i,j,k]*lay_pump[i,1]*t - TDS[i,j,k]*layer_fluxes[i+1,2]*t) / lay_vol[i,1,k]) + 
        RWI_cont[i,1,k]
    }
  }
}
```

## Plot 1: time, layer depth, TDS, simulation df
```{r}
# gather appraoch 
library(tidyr)

# make depth-time array
ld   <- lapply(1:1000, function(x){return(layer_depths_array[1:7,,x])}) # all layer depths in a list
td   <- lapply(1:1000, function(x){return(TDS[1:7,,x])}) # tds as a list of matrices
ldtd <- lapply(1:1000, function(x){return(data.frame(cbind(ld[[x]], td[[x]])))}) # bind

for(i in 1:1000){
  colnames(ldtd[[i]]) <- c("d", "0", "50","100","150","200","250","300","350")
  ldtd[[i]] <- gather(ldtd[[i]], time, tds, -d) # wide to long
  ldtd[[i]]$sim <- i # add simulation
}

df <- bind_rows(ldtd)
df$sim <- as.factor(df$sim)
df$time <- factor(paste0("t = ", df$time), 
                  levels = paste0("t = ", 
                                  c("0", "50","100","150","200","250","300","350")))
df$d <- df$d * 0.3048       # ft to meters

ll <- c(
  't = 0'  ="t = 0 yrs (1960)",
  't = 50' ="t = 50 yrs (2010)",
  't = 100'="t = 100 yrs (2060)",
  't = 150'="t = 150 yrs (2110)",
  't = 200'="t = 200 yrs (2160)",
  't = 250'="t = 250 yrs (2210)",
  't = 300'="t = 300 yrs (2260)",
  't = 350'="t = 350 yrs (2310)"
)



p <- df %>% filter(time %in% c("t = 0", "t = 50", "t = 100", "t = 200", "t = 250", "t = 350") & sim %in% 251:500) %>% 
  ggplot(aes(-d, tds)) + 
  geom_line(aes(color = sim), alpha = 0.2) +
  geom_smooth(se = FALSE, color = "red") + 
  coord_flip(ylim = c(0, 11000)) +
  scale_color_grey() + 
  guides(color = FALSE) + 
  theme_minimal() + 
  facet_wrap(~time, labeller = as_labeller(ll)) +
  scale_y_continuous(breaks = c(1000, 5000, 10000), labels = c('1,000', '5,000', '10,000')) + 
  labs(x = "Depth (m)", y = "TDS (mg/L)")

p

ggsave(p, filename = "results/p_sim.pdf", device = cairo_pdf, height = 5, width = 7)
```


## Plot 2: Compare concentrations of GW, TAW, NDP
```{r Plot 2: Compare concentrations of GW, TAW, NDP}
# initalize matricies
GWP_mean = as.data.frame(matrix(0,7,1))
AW_mean = as.data.frame(matrix(0,7,1))
EC_mean = as.data.frame(matrix(0,7,1))
EC_median = as.data.frame(matrix(0,7,1))
GWP_5 = as.data.frame(matrix(0,7,1))
AW_5 = as.data.frame(matrix(0,7,1))
EC_5 = as.data.frame(matrix(0,7,1))
GWP_95 = as.data.frame(matrix(0,7,1))
AW_95 = as.data.frame(matrix(0,7,1))
EC_95 = as.data.frame(matrix(0,7,1))
for(i in 1:7){
  GWP_mean[i,1] = mean(TDS_GWP[i,1,])
  AW_mean[i,1] = mean(TDS_AW[i,1,])
  EC_mean[i,1] = mean(EC[i,1,])
  EC_median[i,1] = median(EC[i,1,])
  GWP_5[i,1] = quantile(TDS_GWP[i,1,], c(0.05))
  GWP_95[i,1] = quantile(TDS_GWP[i,1,], c(0.95))
  AW_5[i,1] = quantile(TDS_AW[i,1,], c(0.05))
  AW_95[i,1] = quantile(TDS_AW[i,1,], c(0.95))
  EC_5[i,1] = quantile(EC[i,1,], c(0.05))
  EC_95[i,1] = quantile(EC[i,1,], c(0.95))
}

## compile stats for TDS of pumped GW, Total Applied Water (GW+SW), and Evaporative concentration of NDP
EC_dat = data.frame("legend"=rep(c("A","B","C"), each=7),
                    "time" = factor(seq(0,300,50)),
                    "mean" = rbind(GWP_mean, AW_mean, EC_median),
                    "p5" = rbind(GWP_5, AW_5, EC_5),
                    "p95" = rbind(GWP_95, AW_95, EC_95))
colnames(EC_dat) = c("legend","time","mean","p5", "p95")

# plot
p2 <- ggplot(EC_dat, aes(x=time, y=mean, fill=legend)) +
  geom_bar(position=position_dodge(), stat="identity", 
           colour = "black", size=.3,
           alpha = 0.7) +
  geom_errorbar(aes(ymin=EC_dat$p5, ymax=EC_dat$p95),
                size=.3,    # thinner lines
                width=.2,                    # Width of the error bars
                position=position_dodge(.9)) +
  xlab("Time (yrs)") +
  ylab("TDS (mg/L)") +
  scale_y_continuous(labels = comma) +
  scale_fill_viridis_d(name="",
                 breaks=c("A","B","C"),
                 labels=c("Pumped Groundwater", 
                          "Total Applied Water", 
                          "Net Deep Percolation")) +
  theme_minimal() +
  theme(legend.position = c(0.22, 0.75))

p2
# save
ggsave("results/p_ec.pdf", p2, height= 4, width = 7, device = cairo_pdf)
```


# Plot 3: Salt Budget for Water Budget terms and NDP
```{r}
Mtons_to_metric_tons <- function(x){return(x * 907184.74)}

## non:constant mass: extract all GW pumping TDS and convert to mass
gwp <- vector("list", length = 7)

for(i in 1:7){
  gwp[[i]] <- data.frame(TDS_GWP[i, 1, ], t = i, type = "GWP")
}

gwp <- dplyr::bind_rows(gwp)
colnames(gwp) <- c("c","t", "type")
gwp <- gwp %>% 
  mutate(m = Mtons_to_metric_tons(mg_to_Mton(c * annual_fluxes[7,2])))

# constant mass: surface water and allother budget terms

# surface water
swm <- Mtons_to_metric_tons(SW_annual_mass)

# all other budget terms
all_else <- Mtons_to_metric_tons(mg_to_Mton(sum(annual_fluxes[c(2,3,4,5,6,8),4])))

# combine into one df
df1 <- data.frame(m = rep(c(swm, all_else), each = 7), 
                  t = rep(1:7, times = 2), 
                  type = rep(c("surface_water_import", "all_else"), each = 7)) 
df2 <- select(gwp, m, t, type) 
df2 <- df2 %>% 
  group_by(t) %>% 
  summarise(median = median(m), p5 = quantile(m, 0.05), p95 = quantile(m, 0.95))

df1$p5  <- NA
df1$p95 <- NA
df2 <- df2 %>% rename(m = median) %>% mutate(type = "gwp") %>% select(m, t, type, p5, p95)

df3 <- rbind.data.frame(df1, df2)
df3$t <- rep(seq(0,300, 50), times = 3) # format times
df3$type <- factor(rep(c("Surface Water Diversions", "I, C, W, R, L, B", "Pumped Groundwater"), each = 7), # format labels
                   levels = c("I, C, W, R, L, B", "Surface Water Diversions", "Pumped Groundwater"))

# plot
p3 <- ggplot(df3, aes(factor(t), m, fill = type)) +
  geom_bar(position = position_dodge(), stat = "identity", 
           colour = "black", size=.3,
           alpha = 0.7) +
  geom_errorbar(aes(ymin = p5, ymax = p95),
                size=.3,    # thinner lines
                width=.2,   # Width of the error bars
                position=position_dodge(.9)) +
  scale_fill_viridis_d("Water Budget Term", option = "E") +
  theme_minimal() +
  theme(legend.position = c(0.22, 0.75)) +
  labs(x = "Time", y = "Annual Mass (Metric Tons)")

ggsave("results/p_salt_budget.pdf", p3, height= 4, width = 7, device = cairo_pdf)
```


































<!-- ## Table 1: Mass loading into entire GW system from various water budget terms and sources (Mtons) -->
<!-- ```{r Table 1: Mass loading into entire GW system from various water budget terms and sources (Mtons), warning=F, message=F} -->
<!-- # RWI mean, and CIs -->
<!-- RWI_mean = mean(colSums((RWI_cont[1:7,1,]*lay_vol[1:7,1,])/50)) # mg -->
<!-- RWI_5 = quantile(colSums(RWI_cont[1:7,1,]), c(0.05)) -->
<!-- RWI_95 = quantile(colSums(RWI_cont[1:7,1,]), c(0.95)) -->

<!-- # table setup -->
<!-- tstep = 1 # time step of NDP -->
<!-- table1 = data.frame("term" = c("Stream","R","Lake","BI","Sub","SW_D", -->
<!--                                "SI","EC_mean", "EC_5","EC_95",  -->
<!--                                "RWI_mean","RWI_5", "RWI_95"), -->
<!--                     "mass" = c(mg_to_Mton(annual_fluxes[2,4]), mg_to_Mton(annual_fluxes[3,4]), -->
<!--                                mg_to_Mton(annual_fluxes[4,4]), mg_to_Mton(annual_fluxes[5,4]), -->
<!--                                mg_to_Mton(annual_fluxes[6,4]), SW_annual_mass, -->
<!--                                mg_to_Mton(annual_fluxes[8,4]), -->
<!--                                mg_to_Mton(EC_mean[tstep,1] * annual_fluxes[1,2]), -->
<!--                                mg_to_Mton(EC_5[tstep,1] * annual_fluxes[1,2]),  -->
<!--                                mg_to_Mton(EC_95[tstep,1] * annual_fluxes[1,2]), -->
<!--                                mg_to_Mton(RWI_mean),  -->
<!--                                mg_to_Mton(RWI_5), -->
<!--                                mg_to_Mton(RWI_95))) -->
<!-- table1 -->
<!-- ``` -->


<!-- ## Plot 3: Average Vertical Groundwater Pore Velocity Profile -->
<!-- ```{r Plot 3: Average Vertical Groundwater Pore Velocity Profile, warning=F, message=F} -->
<!-- # organize data -->
<!-- plot3data = data.frame(x = layer_velocity_mean[1:7,1],  -->
<!--                   y = as.matrix(layer_depths_mean[1:7,1])) -->
<!-- plot3 = ggplot(plot3data, aes(x = x, y = y)) +  -->
<!--   geom_smooth(method="lm", se= FALSE, color = c("gray40"), formula = y ~ x) + -->
<!--   geom_point(shape=16, cex=3) + -->
<!--   xlab("averge linear pore velocity (ft/yr)") + -->
<!--   ylab("depth below land surface") + -->
<!--   theme_bw() + -->
<!--   ggtitle("Mean Linear Pore Velocity") + -->
<!--   theme(axis.text = element_text(size=16), axis.title = element_text(size=18)) +  -->
<!--   theme(text=element_text(family="Times New Roman", size=18))  -->

<!-- plot3 -->
<!-- ``` -->





<!-- ## Plot 4: Annual mass loading in entire system from RWI and non-RWI sources per time step -->
<!-- ```{r Plot 4: Annual mass loading in entire system from RWI and non-RWI sources per time step} -->
<!-- ## calculate annual mass loading in entire system from RWI and non-RWI sources per time step -->
<!-- table1 = matrix(0,4,7) -->
<!-- for(tstep in 1:7){ -->
<!--   table1[1,tstep] = sum(mg_to_Mton(annual_fluxes[2,4]), mg_to_Mton(annual_fluxes[3,4]), -->
<!--                         mg_to_Mton(annual_fluxes[4,4]), mg_to_Mton(annual_fluxes[5,4]), -->
<!--                         mg_to_Mton(annual_fluxes[6,4]), SW_annual_mass, -->
<!--                         mg_to_Mton(annual_fluxes[8,4])) -->
<!--   table1[2,tstep] = mg_to_Mton(EC_median[tstep,1] * annual_fluxes[1,2])  -->
<!--   table1[3,tstep] = mg_to_Mton(EC_5[tstep,1] * annual_fluxes[1,2]) -->
<!--   table1[4,tstep] = mg_to_Mton(EC_95[tstep,1] * annual_fluxes[1,2]) -->
<!--   #table1[4,tstep] = mg_to_Mton(RWI_mean) -->
<!-- } -->

<!-- table1 = as.data.frame(table1) -->
<!-- rownames(table1) = c('water_budget_terms','evaporite_median', 'evaporite_5', 'evaporite_95') -->
<!-- colnames(table1) = paste0(seq(50,350,50), " yrs") -->
<!-- table1 = t.data.frame(table1) -->

<!-- mass_dat = data.frame("legend"=rep(c("A","B"), each=7), -->
<!--                     "time" = factor(rep(seq(50,350,50), times = 2)), -->
<!--                     "med" = c(table1[,1], table1[,2]), -->
<!--                     "p5" = c(rep(0,7), table1[,3]), -->
<!--                     "p95" = c(rep(0,7), table1[,4])) -->

<!-- Mtons_to_metric_tons <- function(x){return(x * 907184.74)} -->

<!-- mass_dat <- mass_dat %>%  -->
<!--   mutate(med = Mtons_to_metric_tons(med), -->
<!--          p5  = Mtons_to_metric_tons(p5), -->
<!--          p95 = Mtons_to_metric_tons(p95)) -->

<!-- # plot 4: Annual mass loading in entire system from RWI and non-RWI sources per time step -->
<!-- plot4 = ggplot(mass_dat, aes(x=time, y=med, fill=legend)) + -->
<!--   geom_bar(position="dodge", stat="identity",  -->
<!--            colour = "black", size=.3) + -->
<!--   geom_errorbar(aes(ymin=p5, ymax=p95), -->
<!--                 size=.3,    # thinner lines -->
<!--                 width=.2,   # Width of the error bars -->
<!--                 position=position_dodge(.9)) + -->
<!--   xlab("Time (yrs)") + -->
<!--   ylab("mass (Metric Tons)") + -->
<!--   scale_fill_hue(name="", -->
<!--                  breaks=c("A","B"), -->
<!--                  labels=c("I, C, W, R, L, B",  -->
<!--                           "N")) + -->
<!--   #ggtitle("Average Annual Mass Loading in Entire System") + -->
<!--   theme_minimal() +  -->
<!--   scale_y_continuous() -->

<!-- plot4 -->
<!-- ``` -->



<!-- ## Plot 1: Groundwater Salinization Evolution: 0-350 yrs -->
<!-- ```{r Plot 1: Groundwater Salinization Evolution: 0-350 yrs, message=F, warning=F} -->
<!-- # for Plot 1: -->
<!-- TDS_mean = as.data.frame(matrix(0,8,8)) -->
<!-- for(i in 1:8){ -->
<!--   for(j in 1:8){ -->
<!--     TDS_mean[i,j] = mean(TDS[i,j,]) -->
<!--   } -->
<!-- } -->

<!-- sd.top = matrix(0,7,8) -->
<!-- sd.bottom = matrix(0,7,8) -->
<!-- for(d in 1:7){ -->
<!--   for(t in 1:8){ -->
<!--     sd.top[d,t] = TDS_mean[d,t] - sd(TDS[d,t,]) # sd -->
<!--     sd.bottom[d,t] = TDS_mean[d,t] + sd(TDS[d,t,]) # sd -->
<!--   } -->
<!-- } -->

<!-- TDS_5 = matrix(0,7,8) -->
<!-- TDS_95 = matrix(0,7,8) -->
<!-- for(d in 1:7){ -->
<!--   for(t in 1:8){ -->
<!--     TDS_5[d,t] = quantile(TDS[d,t,], c(0.05)) # 5% CI -->
<!--     TDS_95[d,t] = quantile(TDS[d,t,], c(.95))  # 95% CI -->
<!--   } -->
<!-- } -->

<!-- ## calculate layer depths/velocities of simulation means for main plot -->
<!-- layer_depths_mean = as.data.frame(matrix(0,7,1)) -->
<!-- layer_velocity_mean = as.data.frame(matrix(0,7,1)) -->
<!-- for(j in 1:7){ -->
<!--   layer_depths_mean[j,1] = mean(layer_depths_array[j,1,]) * -1 -->
<!--   layer_velocity_mean[j,1] = mean(layer_velocity_array[j,1,]) -->
<!-- } -->

<!-- ## reshape layer_depths_mean for plotting -->
<!-- layer_depths_mean = t.default(layer_depths_mean) -->

<!-- ## trim 8th layer of TDS_mean matrix for plotting -->
<!-- TDS_mean = TDS_mean[1:7,]  -->


<!-- ## data frame of all variables to plot -->
<!-- # pl = data.frame(depth = layer_depths_mean, TDS1 = TDS_mean[,1], TDS2 = TDS_mean[,2], -->
<!-- #                 TDS3 = TDS_mean[,3], TDS4 = TDS_mean[,4], TDS5 = TDS_mean[,5],  -->
<!-- #                 TDS6 = TDS_mean[,6], TDS7 = TDS_mean[,7], TDS8 = TDS_mean[,8], -->
<!-- #                 U1 = TDS_95[,1], L1 = TDS_5[,1], -->
<!-- #                 U2 = TDS_95[,2], L2 = TDS_5[,2], -->
<!-- #                 U3 = TDS_95[,3], L3 = TDS_5[,3], -->
<!-- #                 U4 = TDS_95[,4], L4 = TDS_5[,4], -->
<!-- #                 U5 = TDS_95[,5], L5 = TDS_5[,5], -->
<!-- #                 U6 = TDS_95[,6], L6 = TDS_5[,6], -->
<!-- #                 U7 = TDS_95[,7], L7 = TDS_5[,7], -->
<!-- #                 U8 = TDS_95[,8], L8 = TDS_5[,8])  -->

<!-- # For plotting both models in one graph, see MM_working_MC.Rmd -->
<!-- pl = data.frame(depth = layer_depths_mean, TDS1 = TDS_mean[,1], TDS2 = TDS_mean[,2], -->
<!--                 TDS3 = TDS_mean[,3], TDS4 = TDS_mean[,4], TDS5 = TDS_mean[,5], -->
<!--                 TDS6 = TDS_mean[,6], TDS7 = TDS_mean[,7], TDS8 = TDS_mean[,8], -->
<!--                 U1 = TDS_95[,1], L1 = TDS_5[,1], -->
<!--                 U2 = TDS_95[,2], L2 = TDS_5[,2], -->
<!--                 U3 = TDS_95[,3], L3 = TDS_5[,3], -->
<!--                 U4 = TDS_95[,4], L4 = TDS_5[,4], -->
<!--                 U5 = TDS_95[,5], L5 = TDS_5[,5], -->
<!--                 U6 = TDS_95[,6], L6 = TDS_5[,6], -->
<!--                 U7 = TDS_95[,7], L7 = TDS_5[,7], -->
<!--                 U8 = TDS_95[,8], L8 = TDS_5[,8]) -->

<!-- ymax = 10000 # maximum TDS concentration (mg/L) to display on y axis -->
<!-- ymin = 0 -->

<!-- ## 0 yrs -->
<!-- p1 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS1), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L1, ymax = pl$U1), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("0 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(family="Times New Roman", size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- ## 50 yrs -->
<!-- p2 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS2), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L2, ymax = pl$U2), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("50 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(family="Times New Roman", size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- ## 100 yrs -->
<!-- p3 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS3), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L3, ymax = pl$U3), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("100 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- # for comparison to data -->
<!-- # p3 = ggplot(pl, aes(depth)) + -->
<!-- #   geom_line(aes(y = pl$TDS3), color = 'red') + -->
<!-- #   geom_ribbon(aes(ymin = pl$L3, ymax = pl$U3), alpha=.4) + -->
<!-- #   aes(ymax = ymax) + -->
<!-- #   ylab("TDS (mg/L)") + -->
<!-- #   xlab("depth below land surface (ft.)")  + -->
<!-- #   ggtitle("100 years") + -->
<!-- #   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!-- #   theme(axis.text=element_text(size=14)) + -->
<!-- #   #theme(text=element_text(family="Times New Roman", size = 20)) + -->
<!-- #   coord_flip(ylim = c(100,10000), xlim= c(-1000,-200)) + -->
<!-- #   scale_y_log10(breaks = c(0,10,100,1000,10000,100000), labels = comma(c(0,10,100,1000,10000,100000)), limits = c(10,100000)) -->

<!-- ## 150 yrs -->
<!-- p4 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS4), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L4, ymax = pl$U4), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("150 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(family="Times New Roman", size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- ## 200 yrs -->
<!-- p5 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS5), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L5, ymax = pl$U5), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("200 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(family="Times New Roman", size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- ## 250 yrs -->
<!-- p6 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS6), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L6, ymax = pl$U6), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("250 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(family="Times New Roman", size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- ## 300 yrs -->
<!-- p7 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS7), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L7, ymax = pl$U7), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("300 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(family="Times New Roman", size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- ## 350 yrs -->
<!-- p8 = ggplot(pl, aes(depth)) + -->
<!--   geom_line(aes(y = pl$TDS8), color = 'red') + -->
<!--   geom_ribbon(aes(ymin = pl$L8, ymax = pl$U8), alpha=.4) +  -->
<!--   aes(ymax = ymax, ymin = ymin) + -->
<!--   ylab("TDS (mg/L)") + -->
<!--   xlab("depth below land surface (ft.)")  +  -->
<!--   ggtitle("350 years") +  -->
<!--   geom_hline(yintercept = 1000, linetype = "longdash") + -->
<!--   theme_light() +  -->
<!--   theme(text=element_text(family="Times New Roman", size = 20)) +  -->
<!--   coord_flip() +  -->
<!--   scale_y_continuous(label=comma, breaks = c(1000,5000,10000)) -->

<!-- # save -->
<!-- # for(i in 1:8){ -->
<!-- #   ggsave(get(paste0("p",i)), file = paste0("p",i,".png"), dpi = 300, height = 6, width = 8) -->
<!-- # } -->

<!-- ## arrange p1 - p8 into one plot -->
<!-- #p <- grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol=3, nrow=3, vp=viewport(width=0.95, height=0.95)) -->

<!-- #ggsave("p_no_RWI.PNG", p, height = 8.5, width = 11, units = "in", dpi = 300) -->
<!-- #ggsave("p3_no_RWI.PNG", p3, height = 8.5, width = 11, units = "in", dpi = 300) -->
<!-- ``` -->



